---
title: "model_performance"
author: "Yihan Wang"
date: "2025-02-13"
output: pdf_document
---

```{r}
rm(list = ls())
library(survival)
library(broom)
library(ggplot2)
library(dplyr)
library(caret)
library(pROC)
library(factoextra)
library(timeROC)
library(missRanger)
library(survminer)
library(forcats)
library(tidyr)
library(tibble)
library(penalized)
library(MASS)
library(glmnet)
library(conflicted)
library(gridExtra)
conflict_prefer("select", "dplyr")
conflicts_prefer(dplyr::filter)
set.seed(42)
```

```{r}
source("time-varying-utility.R")
```

```{r}
NACC_full_df_MCI <- read.csv("../preprocessed_data/longitudinal_NACC_MCI.csv")
AIBL_full_df_MCI <- read.csv("../preprocessed_data/longitudinal_AIBL_MCI.csv")

Cog_fts_MCI = c("CDR_SB", "GDS", "MMSE", "memory_recall1", "memory_recall2", "Digit_Span_F", "Digit_Span_B", "animal", "BNT")
demo_fts_MCI = c("age", "sex", "edu","BMI", "blood_pressure_diastolic", "blood_pressure_systolic", "Heart_rate")
basic_fts_MCI = c("APOE4", "age")
index_feature_MCI = c("start", "stop", "event", "ID")
```

```{r}
AIBL_full_df_imputed_MCI<- missranger_impute(AIBL_full_df_MCI, c(c(index_feature_MCI), "Classification"))
```



# internal validation
## time varying cox
### APOE4 + age
```{r}
time_varying_APOE4_NACC_full_results1_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(basic_fts_MCI, index_feature_MCI))], folds = 3, iterations=10)
```



```{r}
time_varying_APOE4_NACC_full_results2_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(basic_fts_MCI, index_feature_MCI))], folds = 3, iterations=10, shorten_visits = TRUE, n_visits=3)
```

```{r}
time_varying_APOE4_NACC_full_results3_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(basic_fts_MCI, index_feature_MCI))], folds = 3, iterations=10, shorten_visits = TRUE, n_visits=2)
```
```{r}
time_varying_APOE4_NACC_full_results4_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(basic_fts_MCI, index_feature_MCI))], folds = 3, iterations=10, shorten_visits = TRUE, n_visits=1)
```

```{r}
plot_auc_trajectory_group(list(time_varying_APOE4_NACC_full_results1_MCI$AUC_Stats, time_varying_APOE4_NACC_full_results2_MCI$AUC_Stats, time_varying_APOE4_NACC_full_results3_MCI$AUC_Stats, time_varying_APOE4_NACC_full_results4_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits", "1-visits"), title_name = " ", output_file = "../results/MCI time-varying AUC internal APOE4.pdf")
```

### all Feature
```{r}
time_varying_all_NACC_results1_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=10)
```



```{r}
time_varying_all_NACC_results2_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=10, shorten_visits = TRUE, n_visits=3)
```

```{r}
time_varying_all_NACC_results3_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=10, shorten_visits = TRUE, n_visits=2)
```

```{r}
time_varying_all_NACC_results4_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=10, shorten_visits = TRUE, n_visits=1)
```

```{r}
plot_auc_trajectory_group(list(time_varying_all_NACC_results1_MCI$AUC_Stats, time_varying_all_NACC_results2_MCI$AUC_Stats, time_varying_all_NACC_results3_MCI$AUC_Stats, time_varying_all_NACC_results4_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits", "1-visits"), title_name = " ", output_file = "../results/MCI time-varying AUC internal all fts.pdf")
```

### Feature Selection - AIC
```{r}
time_varying_all_NACC_AIC_results1_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=10, feature_selection = TRUE)
num_ft_MCI = round(sum(time_varying_all_NACC_AIC_results1_MCI$Significant_Features_Freq$Frequency) / 50)
final_selected_feature_MCI = head(time_varying_all_NACC_AIC_results1_MCI$Significant_Features_Freq$Feature, num_ft_MCI)
```



```{r}
time_varying_all_NACC_AIC_results2_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=10, feature_selection = TRUE, shorten_visits = TRUE, n_visits=3)
```

```{r}
time_varying_all_NACC_AIC_results3_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=10, feature_selection = TRUE, shorten_visits = TRUE, n_visits=2)
```

```{r}
time_varying_all_NACC_AIC_results4_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=10, feature_selection = TRUE, shorten_visits = TRUE, n_visits=1)
```

```{r}
filtered_data_MCI <- time_varying_all_NACC_AIC_results1_MCI$Significant_Features_Freq %>%
  filter(Feature != "frailty(ID)")

ft_plot_MCI <- ggplot(filtered_data_MCI, aes(x = Frequency, y = reorder(Feature, Frequency))) +
  geom_bar(stat = "identity", fill = "#619cff") +
  theme_minimal() +
  labs(title = "Feature Frequency", x = "Frequency", y = "Feature") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid = element_blank()  # Remove all grid lines
  )

# Save the plot as a PDF
ggsave("../results/MCI feature_importance_plot.pdf", plot = ft_plot_MCI, device = "pdf", width = 10, height = 5)

print(ft_plot_MCI)
```

```{r}
plot_auc_trajectory_group(list(time_varying_all_NACC_AIC_results1_MCI$AUC_Stats, time_varying_all_NACC_AIC_results2_MCI$AUC_Stats, time_varying_all_NACC_AIC_results3_MCI$AUC_Stats, time_varying_all_NACC_AIC_results4_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits", "1-visits"), title_name = " ", output_file = "../results/MCI time-varying AUC internal AIC.pdf")
```

## aggregated normal cox
### APOE + Age
```{r}
aggregate_APOE4_NACC_results1_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, c("APOE4"), c("age"), folds = 5, iterations=10)
```



```{r}
aggregate_APOE4_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, c("APOE4"), c("age"), folds = 5, iterations=10, shorten_visits = TRUE, n_visits=3)
```

```{r}
aggregate_APOE4_NACC_results3_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, c("APOE4"), c("age"), folds = 5, iterations=10, shorten_visits = TRUE, n_visits=2)
```

```{r}
plot_auc_trajectory_group(list(aggregate_APOE4_NACC_results1_MCI$AUC_Stats, aggregate_APOE4_NACC_results2_MCI$AUC_Stats, aggregate_APOE4_NACC_results3_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits"), title_name = " ", output_file = "../results/MCI agg-cox AUC internal APOE4.pdf")
```


### all features
```{r}
static_fts_MCI = c("age", "sex", "edu", "APOE4")
dynamic_fts_MCI = c("CDR_SB", "GDS", "MMSE", "memory_recall1", "memory_recall2", "Digit_Span_F", "Digit_Span_B", "animal", "BNT","BMI", "blood_pressure_diastolic", "blood_pressure_systolic", "Heart_rate")


aggregate_all_NACC_results1_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=10)
```



```{r}
aggregate_all_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts, dynamic_fts_MCI, folds = 5, iterations=10, shorten_visits = TRUE, n_visits=3)
```

```{r}
aggregate_all_NACC_results3_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=10, shorten_visits = TRUE, n_visits=2)
```

```{r}
plot_auc_trajectory_group(list(aggregate_all_NACC_results1_MCI$AUC_Stats, aggregate_all_NACC_results2_MCI$AUC_Stats, aggregate_all_NACC_results3_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits"), title_name = " ", output_file = "../results/MCI agg-cox AUC internal all fts.pdf")
```

### Feature Selection - AIC
```{r}
aggregate_fts_NACC_results1_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=10,feature_selection = TRUE)
```

```{r}
num_ft_agg_MCI = round(sum(aggregate_fts_NACC_results1_MCI$Significant_Features_Freq$Frequency) / 50)
final_selected_feature_agg_MCI = head(aggregate_fts_NACC_results1_MCI$Significant_Features_Freq$Feature, num_ft_agg_MCI)
print(final_selected_feature_agg_MCI)
```

```{r}
aggregate_fts_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=10, shorten_visits = TRUE, n_visits=3)
```

```{r}
aggregate_fts_NACC_results3_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=10, shorten_visits = TRUE, n_visits=2)
```

```{r}
plot_auc_trajectory_group(list(aggregate_fts_NACC_results1_MCI$AUC_Stats, aggregate_fts_NACC_results2_MCI$AUC_Stats, aggregate_fts_NACC_results3_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits"), title_name = " ", output_file = "../results/MCI agg-cox AUC internal AIC.pdf")
```

## baseline normal cox (initial visit)
```{r}
NACC_baseline_df_MCI = prepare_for_survival_analysis_n_visits(NACC_full_df_MCI, n = 1)
NACC_baseline_df_MCI$time = NACC_baseline_df_MCI$stop - NACC_baseline_df_MCI$start
```

```{r}
baseline_APOE4_NACC_results_MCI <- do_internal_cv_normal_cox(NACC_baseline_df_MCI[, c("APOE4", "age", "event", "time", "ID")], folds = 5, iterations=10)
```

```{r}
baseline_all_NACC_results_MCI <- do_internal_cv_normal_cox(NACC_baseline_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, c("event", "time", "ID") ))], folds = 5, iterations=10)
```

```{r}
baseline_fts_NACC_results_MCI <- do_internal_cv_normal_cox(NACC_baseline_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, c("event", "time", "ID") ))], folds = 5, iterations=10, feature_selection = TRUE)
```



## baseline normal cox (last visit)
```{r}
NACC_baseline_df2_MCI = prepare_for_survival_analysis_n_visits_backward(NACC_full_df_MCI, n = 1)
NACC_baseline_df2_MCI$time = NACC_baseline_df2_MCI$stop - NACC_baseline_df2_MCI$start
```

```{r}
baseline_APOE4_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_baseline_df2_MCI[, c("APOE4", "age", "event", "time", "ID")], folds = 5, iterations=10)
```

```{r}
baseline_all_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_baseline_df2_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, c("event", "time", "ID") ))], folds = 5, iterations=10)
```

```{r}
baseline_fts_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_baseline_df2_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, c("event", "time", "ID") ))], folds = 5, iterations=10, feature_selection = TRUE)
```



##########################################################################
# External Validation
## time-varying
```{r}
print(final_selected_feature_MCI)
```

```{r}
selected_features_MCI = c("memory_recall2", "GDS", "edu", "age", "animal", "MMSE")
time_varying_external_results1_MCI <- do_external_time_varying_cox(NACC_full_df_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], AIBL_full_df_imputed_MCI[, unique(c(selected_features_MCI, index_feature_MCI))])
```

```{r}
create_forest_plot(time_varying_external_results1_MCI$model_summary, "Forest Plot for Time-Varying Cox Model", "../results/MCI forest_plot.pdf")
```



```{r}
time_varying_external_results2_MCI <- do_external_time_varying_cox(NACC_full_df_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], AIBL_full_df_imputed_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], shorten_visits = TRUE, n_visits = 3)
```

```{r}
time_varying_external_results3_MCI <- do_external_time_varying_cox(NACC_full_df_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], AIBL_full_df_imputed_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], shorten_visits = TRUE, n_visits = 2)
```

```{r}
time_varying_external_results4_MCI <- do_external_time_varying_cox(NACC_full_df_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], AIBL_full_df_imputed_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], shorten_visits = TRUE, n_visits = 1)
```


```{r}
plot_auc_trajectory_single(list(time_varying_external_results1_MCI$AUC_Stats, time_varying_external_results2_MCI$AUC_Stats, time_varying_external_results3_MCI$AUC_Stats, time_varying_external_results4_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits", "1-visits"), title_name = " ", output_file = "../results/MCI time-varying AUC external.pdf")
```

## aggregated normal cox

```{r}
static_fts2_MCI = c("edu", "APOE4","age")
# dynamic_fts2 = c("CDR_SB", "MMSE",  "memory_recall2", "age",  "animal", "Heart_rate", "blood_pressure_diastolic", "GDS")
dynamic_fts2_MCI = c("CDR_SB", "MMSE",  "memory_recall2", "animal", "GDS")

aggregate_fts_AIBL_results1_MCI <- do_external_normal_cox(NACC_full_df_MCI, AIBL_full_df_imputed_MCI, static_fts2_MCI, dynamic_fts2_MCI)
```



```{r}
aggregate_fts_AIBL_results2_MCI <- do_external_normal_cox(NACC_full_df_MCI, AIBL_full_df_imputed_MCI, static_fts2_MCI, dynamic_fts2_MCI, shorten_visits = TRUE, n_visits = 3)
```

```{r}
aggregate_fts_AIBL_results3_MCI <- do_external_normal_cox(NACC_full_df_MCI, AIBL_full_df_imputed_MCI, static_fts2_MCI, dynamic_fts2_MCI, shorten_visits = TRUE, n_visits = 2)
```

```{r}
plot_auc_trajectory_single(list(aggregate_fts_AIBL_results1_MCI$AUC_Stats, aggregate_fts_AIBL_results2_MCI$AUC_Stats, aggregate_fts_AIBL_results3_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits"), title_name = " ", output_file = "../results/MCI aggre-cox AUC external.pdf")
```


## baseline normal cox (initial visit)
```{r}
AIBL_baseline_df_MCI = prepare_for_survival_analysis_n_visits(AIBL_full_df_imputed_MCI, n = 1)
AIBL_baseline_df_MCI$time = AIBL_baseline_df_MCI$stop - AIBL_baseline_df_MCI$start

```


```{r}
baseline_AIBL_results1_MCI <- do_external_normal_cox(NACC_baseline_df_MCI[, unique(c(selected_features_MCI, c("event", "time", "ID") ))], AIBL_baseline_df_MCI[, unique(c(selected_features_MCI, c("event", "time", "ID") ))])
```



```{r}
plot_auc_trajectory_group(list(baseline_all_NACC_results_MCI$AUC_Stats, baseline_APOE4_NACC_results_MCI$AUC_Stats, baseline_fts_NACC_results_MCI$AUC_Stats), c("full features", "APOE+age", "selected features"), validation_auc_stats = baseline_AIBL_results1_MCI$AUC_Stats, validation_label = "External (selected features)", title_name = " ", output_file = "../results/MCI baseline_cox_init AUC.pdf")

```

## baseline normal cox (last visit)
```{r}
AIBL_baseline_df2_MCI = prepare_for_survival_analysis_n_visits_backward(AIBL_full_df_imputed_MCI, n = 1)
AIBL_baseline_df2_MCI$time = AIBL_baseline_df2_MCI$stop - AIBL_baseline_df2_MCI$start
```


```{r}
baseline_AIBL_results2_MCI <- do_external_normal_cox(NACC_baseline_df2_MCI[, unique(c(selected_features_MCI, c("event", "time", "ID") ))], AIBL_baseline_df2_MCI[, unique(c(selected_features_MCI, c("event", "time", "ID") ))])
```



```{r}
plot_auc_trajectory_group(list(baseline_all_NACC_results2_MCI$AUC_Stats, baseline_APOE4_NACC_results2_MCI$AUC_Stats, baseline_fts_NACC_results2_MCI$AUC_Stats), c("full features", "APOE+age", "selected features"), validation_auc_stats = baseline_AIBL_results2_MCI$AUC_Stats, validation_label = "External (selected features)", title_name = " ", output_file = "../results/MCI baseline_cox_last AUC.pdf")

```



# final plots
```{r}
boxplot_c_index_df_MCI <- data.frame(model = character(), feature_set = character(), value = numeric(), stringsAsFactors = FALSE)


boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "baseline-cox[init]", "APOE4+age (internal-NACC)", baseline_APOE4_NACC_results_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "baseline-cox[last]", "APOE4+age (internal-NACC)", baseline_APOE4_NACC_results2_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "aggregated-cox", "APOE4+age (internal-NACC)", aggregate_APOE4_NACC_results1_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "time-cox(FADRS-MCI)", "APOE4+age (internal-NACC)", time_varying_APOE4_NACC_full_results1_MCI$C_Index_List)

boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "baseline-cox[init]", "full features (internal-NACC)", baseline_all_NACC_results_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "baseline-cox[last]", "full features (internal-NACC)", baseline_all_NACC_results2_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "aggregated-cox", "full features (internal-NACC)", aggregate_all_NACC_results1_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "time-cox(FADRS-MCI)", "full features (internal-NACC)", time_varying_all_NACC_results1_MCI$C_Index_List)

boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "baseline-cox[init]", "AIC feature selection (internal-NACC)", baseline_fts_NACC_results_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "baseline-cox[last]", "AIC feature selection (internal-NACC)", baseline_fts_NACC_results2_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "aggregated-cox", "AIC feature selection (internal-NACC)", aggregate_fts_NACC_results1_MCI$C_Index_List)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "time-cox(FADRS-MCI)", "AIC feature selection (internal-NACC)", time_varying_all_NACC_AIC_results1_MCI$C_Index_List)


boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "baseline-cox[init]", "selected feature set (external-AIBL)", baseline_AIBL_results1_MCI$C_Index)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "baseline-cox[last]", "selected feature set (external-AIBL)", baseline_AIBL_results2_MCI$C_Index)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "aggregated-cox", "selected feature set (external-AIBL)", aggregate_fts_AIBL_results1_MCI$C_Index)
boxplot_c_index_df_MCI = add_data(boxplot_c_index_df_MCI, "time-cox(FADRS-MCI)", "selected feature set (external-AIBL)", time_varying_external_results1_MCI$C_Index)

boxplot_c_index_df_MCI$model <- factor(boxplot_c_index_df_MCI$model, levels = c("baseline-cox[init]", "baseline-cox[last]", "aggregated-cox", "time-cox(FADRS-MCI)"))
boxplot_c_index_df_MCI$feature_set <- factor(boxplot_c_index_df_MCI$feature_set, levels = c("APOE4+age (internal-NACC)", "full features (internal-NACC)", "AIC feature selection (internal-NACC)", "selected feature set (external-AIBL)"))

boxplot_c_index_df_MCI <- boxplot_c_index_df_MCI %>% filter(is.finite(value))
```


```{r}
plot_boxplot(boxplot_c_index_df_MCI, "C-index", "../results/MCI boxplot_c_index.pdf")

```


```{r}
boxplot_auc_index_df_MCI <- data.frame(model = character(), feature_set = character(), value = numeric(), stringsAsFactors = FALSE)

boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "baseline-cox[init]", "APOE4+age (internal-NACC)", na.omit(baseline_APOE4_NACC_results_MCI$AUC_list))

boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "baseline-cox[init]", "full features (internal-NACC)", na.omit(baseline_all_NACC_results_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "baseline-cox[init]", "AIC feature selection (internal-NACC)", na.omit(baseline_fts_NACC_results_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "baseline-cox[init]", "selected feature set (external-AIBL)", na.omit(baseline_AIBL_results1_MCI$AUC_list))


boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "baseline-cox[last]", "APOE4+age (internal-NACC)", na.omit(baseline_APOE4_NACC_results2_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "baseline-cox[last]", "full features (internal-NACC)", na.omit(baseline_all_NACC_results2_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "baseline-cox[last]", "AIC feature selection (internal-NACC)", na.omit(baseline_fts_NACC_results2_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "baseline-cox[last]", "selected feature set (external-AIBL)", na.omit(baseline_AIBL_results2_MCI$AUC_list))

boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "aggregated-cox", "APOE4+age (internal-NACC)", na.omit(aggregate_APOE4_NACC_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "aggregated-cox", "full features (internal-NACC)", na.omit(aggregate_all_NACC_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "aggregated-cox", "AIC feature selection (internal-NACC)", na.omit(aggregate_fts_NACC_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "aggregated-cox", "selected feature set (external-AIBL)", na.omit(aggregate_fts_AIBL_results1_MCI$AUC_list))

boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "time-cox(FADRS-MCI)", "APOE4+age (internal-NACC)", na.omit(time_varying_APOE4_NACC_full_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "time-cox(FADRS-MCI)", "full features (internal-NACC)", na.omit(time_varying_all_NACC_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "time-cox(FADRS-MCI)", "AIC feature selection (internal-NACC)", na.omit(time_varying_all_NACC_AIC_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI = add_data(boxplot_auc_index_df_MCI, "time-cox(FADRS-MCI)", "selected feature set (external-AIBL)", na.omit(time_varying_external_results1_MCI$AUC_list))

boxplot_auc_index_df_MCI$model <- factor(boxplot_auc_index_df_MCI$model, levels = c("baseline-cox[init]", "baseline-cox[last]", "aggregated-cox", "time-cox(FADRS-MCI)"))
boxplot_auc_index_df_MCI$feature_set <- factor(boxplot_auc_index_df_MCI$feature_set, levels = c("APOE4+age (internal-NACC)", "full features (internal-NACC)", "AIC feature selection (internal-NACC)", "selected feature set (external-AIBL)"))

boxplot_auc_index_df_MCI <- boxplot_auc_index_df_MCI %>% filter(is.finite(value))
```

```{r}
plot_boxplot(boxplot_auc_index_df_MCI, "AUC", "../results/MCI boxplot_auc.pdf")

```




####### clustering

```{r}
risk_group_train_MCI <- time_varying_external_results1_MCI$risk_train %>%
  mutate(risk_group = case_when(
    last_risk_score < quantile(last_risk_score, 0.5) ~ "Low Risk",
    # last_risk_score < quantile(last_risk_score, 0.66) ~ "Medium Risk",
    TRUE ~ "High Risk"
  ))
analyze_risk_groups(NACC_full_df_MCI, risk_group_train_MCI)
```


```{r}
risk_group_means_train_MCI <- sapply(unique(risk_group_train_MCI$risk_group), function(group) {
  group_ids <- risk_group_train_MCI$ID[risk_group_train_MCI$risk_group == group]
  subset_results <- time_varying_external_results1_MCI$risk_train[time_varying_external_results1_MCI$risk_train$ID %in% group_ids, ]
  mean(subset_results$last_risk_score, na.rm = TRUE)
})
names(risk_group_means_train_MCI) <- unique(risk_group_train_MCI$risk_group)
risk_group_means_train_MCI
```


```{r}
risk_group_test_MCI <- time_varying_external_results1_MCI$risk_test %>%
  mutate(
    risk_group = sapply(last_risk_score, function(score) {
      differences <- abs(risk_group_means_train_MCI - score)
      names(differences)[which.min(differences)]
    })
  )

analyze_risk_groups(AIBL_full_df_imputed_MCI, risk_group_test_MCI)
```

```{r}
selected_features_trajectories_MCI = c("memory_recall2", "animal", "MMSE", "GDS")
centroid_trajectory_test_MCI = plot_centroid_trajectories(AIBL_full_df_imputed_MCI, risk_group_test_MCI, selected_features_trajectories_MCI, "../results/MCI feature_trajectories_AIBL.pdf")
```



```{r}
risk_group_means_test_MCI <- sapply(unique(risk_group_test_MCI$risk_group), function(group) {
  group_ids <- risk_group_test_MCI$ID[risk_group_test_MCI$risk_group == group]
  subset_results <- time_varying_external_results1_MCI$risk_test[time_varying_external_results1_MCI$risk_test$ID %in% group_ids, ]
  mean(subset_results$last_risk_score, na.rm = TRUE)
})
names(risk_group_means_test_MCI) <- unique(risk_group_test_MCI$risk_group)
risk_group_means_test_MCI
```

```{r}
risk_group_test_MCI

```
