 ---
title: "model_performance"
author: "Yihan Wang"
date: "2025-02-13"
output: pdf_document
---

```{r}
# rm(list = ls())
library(survival)
library(broom)
library(ggplot2)
library(dplyr)
library(caret)
library(pROC)
library(factoextra)
library(timeROC)
library(missRanger)
library(survminer)
library(forcats)
library(tidyr)
library(tibble)
library(penalized)
library(MASS)
library(glmnet)
library(conflicted)
library(gridExtra)
conflict_prefer("select", "dplyr")
conflicts_prefer(dplyr::filter)
set.seed(42)
```

```{r}
source("time-varying-utility.R")
```

```{r}
NACC_full_df_MCI <- read.csv("../preprocessed_data/longitudinal_NACC_MCI.csv")
AIBL_full_df_MCI <- read.csv("../preprocessed_data/longitudinal_AIBL_MCI.csv")

Cog_fts_MCI = c("CDR_SB", "GDS", "MMSE", "memory_recall1", "memory_recall2", "Digit_Span_F", "Digit_Span_B", "animal", "BNT")
demo_fts_MCI = c("age", "sex", "edu","BMI", "blood_pressure_diastolic", "blood_pressure_systolic", "Heart_rate")
basic_fts_MCI = c("APOE4", "age")
index_feature_MCI = c("start", "stop", "event", "ID")
```

```{r}
AIBL_full_df_imputed_MCI<- missranger_impute(AIBL_full_df_MCI, c(c(index_feature_MCI), "Classification"))
```



# internal validation
## time varying cox
### APOE4 + age
```{r}
time_varying_APOE4_NACC_full_results1_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(basic_fts_MCI, index_feature_MCI))], folds = 3, iterations=50)
```



```{r}
time_varying_APOE4_NACC_full_results2_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(basic_fts_MCI, index_feature_MCI))], folds = 3, iterations=50, shorten_visits = TRUE, n_visits=3)
```

```{r}
time_varying_APOE4_NACC_full_results3_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(basic_fts_MCI, index_feature_MCI))], folds = 3, iterations=50, shorten_visits = TRUE, n_visits=2)
```
```{r}
time_varying_APOE4_NACC_full_results4_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(basic_fts_MCI, index_feature_MCI))], folds = 3, iterations=50, shorten_visits = TRUE, n_visits=1)
```

```{r}
plot_auc_trajectory_group(list(time_varying_APOE4_NACC_full_results1_MCI$AUC_Stats, time_varying_APOE4_NACC_full_results2_MCI$AUC_Stats, time_varying_APOE4_NACC_full_results3_MCI$AUC_Stats, time_varying_APOE4_NACC_full_results4_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits", "1-visits"), title_name = " ", output_file = "../results/MCI time-varying AUC internal APOE4.pdf")
```

### all Feature
```{r}
time_varying_all_NACC_results1_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=50)
```



```{r}
time_varying_all_NACC_results2_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=50, shorten_visits = TRUE, n_visits=3)
```

```{r}
time_varying_all_NACC_results3_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=50, shorten_visits = TRUE, n_visits=2)
```

```{r}
time_varying_all_NACC_results4_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=50, shorten_visits = TRUE, n_visits=1)
```

```{r}
plot_auc_trajectory_group(list(time_varying_all_NACC_results1_MCI$AUC_Stats, time_varying_all_NACC_results2_MCI$AUC_Stats, time_varying_all_NACC_results3_MCI$AUC_Stats, time_varying_all_NACC_results4_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits", "1-visits"), title_name = " ", output_file = "../results/MCI time-varying AUC internal all fts.pdf")
```

### Feature Selection - AIC
```{r}
# Internal cross-validation with AIC feature selection
time_varying_all_NACC_AIC_results1_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=50, feature_selection = TRUE)

# Train final model on full NACC dataset with AIC selection
NACC_train_data_MCI <- NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))]
train_predictors_MCI <- setdiff(names(NACC_train_data_MCI), c("start", "stop", "event", "ID"))

# Scale and remove high correlation
data_preprocess_MCI <- remove_high_corr_scales(NACC_train_data_MCI, NACC_train_data_MCI, cols_consider = train_predictors_MCI, threshold = 0.9)
train_data_scaled_MCI <- data_preprocess_MCI$train
train_predictors_scaled_MCI <- setdiff(names(train_data_scaled_MCI), c("start", "stop", "event", "ID"))

# Fit initial model with all features (without frailty for stepAIC)
formula_initial_MCI <- as.formula(paste("Surv(start, stop, event) ~", paste(train_predictors_scaled_MCI, collapse = " + ")))
cox_model_initial_MCI <- coxph(formula_initial_MCI, data = train_data_scaled_MCI, ties = "efron", control = coxph.control(iter.max = 100, eps = 1e-05))

# Apply AIC-based stepwise selection
stepwise_model_MCI <- stepAIC(cox_model_initial_MCI, direction = "both", trace = FALSE)
final_selected_feature_MCI <- setdiff(all.vars(formula(stepwise_model_MCI))[-1], c("start", "stop", "event", "ID"))

cat("Features selected by AIC on full NACC cohort (MCI):n")
print(final_selected_feature_MCI)
```



```{r}
time_varying_all_NACC_AIC_results2_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=50, feature_selection = TRUE, shorten_visits = TRUE, n_visits=3)
```

```{r}
time_varying_all_NACC_AIC_results3_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=50, feature_selection = TRUE, shorten_visits = TRUE, n_visits=2)
```

```{r}
time_varying_all_NACC_AIC_results4_MCI <- do_internal_cv_time_varying_cox(NACC_full_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, index_feature_MCI))], folds = 5, iterations=50, feature_selection = TRUE, shorten_visits = TRUE, n_visits=1)
```

```{r}
# filtered_data_MCI <- time_varying_all_NACC_AIC_results1_MCI$Significant_Features_Freq %>%
#   filter(Feature != "frailty(ID)")
# 
# ft_plot_MCI <- ggplot(filtered_data_MCI, aes(x = Frequency, y = reorder(Feature, Frequency))) +
#   geom_bar(stat = "identity", fill = "#9bbf8a") +
#   theme_minimal() +
#   labs(title = "Feature Frequency", x = "Frequency", y = "Feature") +
#   theme(
#     plot.title = element_text(hjust = 0.5, size = 14),
#     axis.title = element_text(size = 12),
#     axis.text = element_text(size = 10),
#     panel.grid = element_blank()  # Remove all grid lines
#   )
# 
# # Save the plot as a PDF
# ggsave("../results/MCI feature_importance_plot.pdf", plot = ft_plot_MCI, device = "pdf", width = 10, height = 5)
# 
# print(ft_plot_MCI)
# define rename map
feature_rename_MCI <- c(
  "CDR_SB"                    = "CDR Sum of Boxes",
  "MMSE"                      = "MMSE Score",
  "memory_recall2"            = "LM-II",
  "animal"                    = "Animal Fluency Test",
  # "GDS"                       = "Geriatric Depression Scale",
  # "APOE4"                     = "APOE ε4 Status",
  "age"                       = "Age",
  "edu"                       = "Education",
  "sex"                       = "Sex",
  "BMI"                       = "BMI",
  "blood_pressure_diastolic"  = "BP(Diastolic)",
  "blood_pressure_systolic"   = "BP(Systolic)",
  "Heart_rate"                = "Heart Rate",
  "memory_recall1"            = "LM-I",
  "Digit_Span_F"              = "Digit Span Forward",
  "Digit_Span_B"              = "Digit Span Backward",
  "BNT"                       = "Boston Naming Test"
)

filtered_data_MCI <- time_varying_all_NACC_AIC_results1_MCI$Significant_Features_Freq %>%
  filter(Feature != "frailty(ID)") %>%
  mutate(
    Feature_Label = ifelse(
      Feature %in% names(feature_rename_MCI),
      feature_rename_MCI[Feature],
      Feature                          # keep original if not in rename map
    ),
    Feature_Label = factor(Feature_Label,
                           levels = Feature_Label[order(Frequency)])
  )

ft_plot_MCI <- ggplot(filtered_data_MCI,
    aes(x = Frequency, y = Feature_Label)) +
  geom_bar(stat = "identity", fill = "#9bbf8a",
           width = 0.7) +
  labs(
    x = "Frequency",
    y = NULL
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.background  = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line.y      = element_blank(),
    axis.ticks.y     = element_blank(),
    axis.title.x     = element_text(size = 14, face = "bold"),
    axis.text.x      = element_text(size = 10),
    axis.text.y      = element_text(size = 11)
  )

ggsave("../results/MCI_panel_C_feature_importance_plot.pdf",
       plot   = ft_plot_MCI,
       device = "pdf",
       width  = 7,
       height = 4)
       # height = max(3, nrow(filtered_data_MCI) * 0.4 + 1))

print(ft_plot_MCI)
```

```{r}
plot_auc_trajectory_group(list(time_varying_all_NACC_AIC_results1_MCI$AUC_Stats, time_varying_all_NACC_AIC_results2_MCI$AUC_Stats, time_varying_all_NACC_AIC_results3_MCI$AUC_Stats, time_varying_all_NACC_AIC_results4_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits", "1-visits"), title_name = " ", output_file = "../results/MCI time-varying AUC internal AIC.pdf")
```

## aggregated normal cox
### APOE + Age
```{r}
aggregate_APOE4_NACC_results1_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, c("APOE4","age"),c(),folds = 5, iterations=50)
```



```{r}
aggregate_APOE4_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, c("APOE4","age"), c(),  folds = 5, iterations=50, shorten_visits = TRUE, n_visits=3)
```

```{r}
aggregate_APOE4_NACC_results3_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, c("APOE4","age"),c(), folds = 5, iterations=50, shorten_visits = TRUE, n_visits=2)
```

```{r}
plot_auc_trajectory_group(list(aggregate_APOE4_NACC_results1_MCI$AUC_Stats, aggregate_APOE4_NACC_results2_MCI$AUC_Stats, aggregate_APOE4_NACC_results3_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits"), title_name = " ", output_file = "../results/MCI agg-cox AUC internal APOE4.pdf")
```


### all features
```{r}
static_fts_MCI = c("age", "sex", "edu", "APOE4")
dynamic_fts_MCI = c("CDR_SB", "GDS", "MMSE", "memory_recall1", "memory_recall2", "Digit_Span_F", "Digit_Span_B", "animal", "BNT","BMI", "blood_pressure_diastolic", "blood_pressure_systolic", "Heart_rate")


aggregate_all_NACC_results1_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=50)
```



```{r}
aggregate_all_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=50, shorten_visits = TRUE, n_visits=3)
```

```{r}
aggregate_all_NACC_results3_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=50, shorten_visits = TRUE, n_visits=2)
```

```{r}
plot_auc_trajectory_group(list(aggregate_all_NACC_results1_MCI$AUC_Stats, aggregate_all_NACC_results2_MCI$AUC_Stats, aggregate_all_NACC_results3_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits"), title_name = " ", output_file = "../results/MCI agg-cox AUC internal all fts.pdf")
```

### Feature Selection - AIC
```{r}
aggregate_fts_NACC_results1_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=50,feature_selection = TRUE)
```

```{r}
num_ft_agg_MCI = round(sum(aggregate_fts_NACC_results1_MCI$Significant_Features_Freq$Frequency) / 50)
final_selected_feature_agg_MCI = head(aggregate_fts_NACC_results1_MCI$Significant_Features_Freq$Feature, num_ft_agg_MCI)
print(final_selected_feature_agg_MCI)
```

```{r}
aggregate_fts_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=50, shorten_visits = TRUE, n_visits=3,feature_selection = TRUE)
```

```{r}
aggregate_fts_NACC_results3_MCI <- do_internal_cv_normal_cox(NACC_full_df_MCI, static_fts_MCI, dynamic_fts_MCI, folds = 5, iterations=50, shorten_visits = TRUE, n_visits=2,feature_selection = TRUE)
```

```{r}
plot_auc_trajectory_group(list(aggregate_fts_NACC_results1_MCI$AUC_Stats, aggregate_fts_NACC_results2_MCI$AUC_Stats, aggregate_fts_NACC_results3_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits"), title_name = " ", output_file = "../results/MCI agg-cox AUC internal AIC.pdf")
```

## baseline normal cox (initial visit)
```{r}
NACC_baseline_df_MCI = prepare_for_survival_analysis_n_visits(NACC_full_df_MCI, n = 1)
NACC_baseline_df_MCI$time = NACC_baseline_df_MCI$stop - NACC_baseline_df_MCI$start
```

```{r}
baseline_APOE4_NACC_results_MCI <- do_internal_cv_normal_cox(NACC_baseline_df_MCI[, c("APOE4", "age", "event", "time", "ID")], folds = 5, iterations=50)
```

```{r}
baseline_all_NACC_results_MCI <- do_internal_cv_normal_cox(NACC_baseline_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, c("event", "time", "ID") ))], folds = 5, iterations=50)
```

```{r}
baseline_fts_NACC_results_MCI <- do_internal_cv_normal_cox(NACC_baseline_df_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, c("event", "time", "ID") ))], folds = 5, iterations=10, feature_selection = TRUE)
```



## baseline normal cox (last visit)
```{r}
NACC_baseline_df2_MCI = prepare_for_survival_analysis_n_visits_backward(NACC_full_df_MCI, n = 1)
NACC_baseline_df2_MCI$time = NACC_baseline_df2_MCI$stop - NACC_baseline_df2_MCI$start
```

```{r}
baseline_APOE4_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_baseline_df2_MCI[, c("APOE4", "age", "event", "time", "ID")], folds = 5, iterations=50)
```

```{r}
baseline_all_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_baseline_df2_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, c("event", "time", "ID") ))], folds = 5, iterations=50)
```

```{r}
baseline_fts_NACC_results2_MCI <- do_internal_cv_normal_cox(NACC_baseline_df2_MCI[, unique(c(Cog_fts_MCI, demo_fts_MCI, basic_fts_MCI, c("event", "time", "ID") ))], folds = 5, iterations=50, feature_selection = TRUE)
```



##########################################################################
# External Validation
## time-varying
```{r}
print(final_selected_feature_MCI)
```

```{r}
# Use AIC-selected features directly (no manual selection)
selected_features_MCI <- final_selected_feature_MCI
time_varying_external_results1_MCI <- do_external_time_varying_cox(NACC_full_df_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], AIBL_full_df_imputed_MCI[, unique(c(selected_features_MCI, index_feature_MCI))])
```

```{r}
# create_forest_plot(time_varying_external_results1_MCI$model_summary, "Forest Plot for Time-Varying Cox Model", "../results/MCI forest_plot.pdf")
```



```{r}
time_varying_external_results2_MCI <- do_external_time_varying_cox(NACC_full_df_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], AIBL_full_df_imputed_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], shorten_visits = TRUE, n_visits = 3)
```

```{r}
time_varying_external_results3_MCI <- do_external_time_varying_cox(NACC_full_df_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], AIBL_full_df_imputed_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], shorten_visits = TRUE, n_visits = 2)
```

```{r}
time_varying_external_results4_MCI <- do_external_time_varying_cox(NACC_full_df_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], AIBL_full_df_imputed_MCI[, unique(c(selected_features_MCI, index_feature_MCI))], shorten_visits = TRUE, n_visits = 1)
```


```{r}
plot_auc_trajectory_single(list(time_varying_external_results1_MCI$AUC_Stats, time_varying_external_results2_MCI$AUC_Stats, time_varying_external_results3_MCI$AUC_Stats, time_varying_external_results4_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits", "1-visits"), title_name = " ", output_file = "../results/MCI time-varying AUC external.pdf")
```

## aggregated normal cox

```{r}
static_fts2_MCI = c("edu", "APOE4","age")
# dynamic_fts2 = c("CDR_SB", "MMSE",  "memory_recall2", "age",  "animal", "Heart_rate", "blood_pressure_diastolic", "GDS")
dynamic_fts2_MCI = c("CDR_SB", "MMSE",  "memory_recall2", "animal", "GDS")

aggregate_fts_AIBL_results1_MCI <- do_external_normal_cox(NACC_full_df_MCI, AIBL_full_df_imputed_MCI, static_fts2_MCI, dynamic_fts2_MCI)
```



```{r}
aggregate_fts_AIBL_results2_MCI <- do_external_normal_cox(NACC_full_df_MCI, AIBL_full_df_imputed_MCI, static_fts2_MCI, dynamic_fts2_MCI, shorten_visits = TRUE, n_visits = 3)
```

```{r}
aggregate_fts_AIBL_results3_MCI <- do_external_normal_cox(NACC_full_df_MCI, AIBL_full_df_imputed_MCI, static_fts2_MCI, dynamic_fts2_MCI, shorten_visits = TRUE, n_visits = 2)
```

```{r}
plot_auc_trajectory_single(list(aggregate_fts_AIBL_results1_MCI$AUC_Stats, aggregate_fts_AIBL_results2_MCI$AUC_Stats, aggregate_fts_AIBL_results3_MCI$AUC_Stats), c("full visits", "3-visits", "2-visits"), title_name = " ", output_file = "../results/MCI aggre-cox AUC external.pdf")
```


## baseline normal cox (initial visit)
```{r}
AIBL_baseline_df_MCI = prepare_for_survival_analysis_n_visits(AIBL_full_df_imputed_MCI, n = 1)
AIBL_baseline_df_MCI$time = AIBL_baseline_df_MCI$stop - AIBL_baseline_df_MCI$start

```


```{r}
baseline_AIBL_results1_MCI <- do_external_normal_cox(NACC_baseline_df_MCI[, unique(c(selected_features_MCI, c("event", "time", "ID") ))], AIBL_baseline_df_MCI[, unique(c(selected_features_MCI, c("event", "time", "ID") ))])
```



```{r}
plot_auc_trajectory_group(list(baseline_all_NACC_results_MCI$AUC_Stats, baseline_APOE4_NACC_results_MCI$AUC_Stats, baseline_fts_NACC_results_MCI$AUC_Stats), c("full features", "APOE+age", "selected features"), validation_auc_stats = baseline_AIBL_results1_MCI$AUC_Stats, validation_label = "External (selected features)", title_name = " ", output_file = "../results/MCI baseline_cox_init AUC.pdf")

```

## baseline normal cox (last visit)
```{r}
AIBL_baseline_df2_MCI = prepare_for_survival_analysis_n_visits_backward(AIBL_full_df_imputed_MCI, n = 1)
AIBL_baseline_df2_MCI$time = AIBL_baseline_df2_MCI$stop - AIBL_baseline_df2_MCI$start
```


```{r}
baseline_AIBL_results2_MCI <- do_external_normal_cox(NACC_baseline_df2_MCI[, unique(c(selected_features_MCI, c("event", "time", "ID") ))], AIBL_baseline_df2_MCI[, unique(c(selected_features_MCI, c("event", "time", "ID") ))])
```



```{r}
plot_auc_trajectory_group(list(baseline_all_NACC_results2_MCI$AUC_Stats, baseline_APOE4_NACC_results2_MCI$AUC_Stats, baseline_fts_NACC_results2_MCI$AUC_Stats), c("full features", "APOE+age", "selected features"), validation_auc_stats = baseline_AIBL_results2_MCI$AUC_Stats, validation_label = "External (selected features)", title_name = " ", output_file = "../results/MCI baseline_cox_last AUC.pdf")

```



# final plots
```{r}
# ============================================================
# PANEL A: AUC Boxplot — model comparison
# x-axis: APOE4+age | full features | AIC selected | external (AIBL)
# fill: model type
# ============================================================
# ============================================================
# REBUILD AUC BOXPLOT DATAFRAME - MCI
# ============================================================

boxplot_auc_index_df_MCI <- data.frame(
  model = character(), feature_set = character(),
  value = numeric(), stringsAsFactors = FALSE
)

# baseline-cox[init]
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "baseline-cox[init]",
  "APOE4+age (internal-NACC)",             na.omit(baseline_APOE4_NACC_results_MCI$AUC_list))
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "baseline-cox[init]",
  "full features (internal-NACC)",          na.omit(baseline_all_NACC_results_MCI$AUC_list))
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "baseline-cox[init]",
  "AIC feature selection (internal-NACC)",  na.omit(baseline_fts_NACC_results_MCI$AUC_list))

# baseline-cox[last]
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "baseline-cox[last]",
  "APOE4+age (internal-NACC)",             na.omit(baseline_APOE4_NACC_results2_MCI$AUC_list))
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "baseline-cox[last]",
  "full features (internal-NACC)",          na.omit(baseline_all_NACC_results2_MCI$AUC_list))
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "baseline-cox[last]",
  "AIC feature selection (internal-NACC)",  na.omit(baseline_fts_NACC_results2_MCI$AUC_list))

# aggregated-cox
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "aggregated-cox",
  "APOE4+age (internal-NACC)",             na.omit(aggregate_APOE4_NACC_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "aggregated-cox",
  "full features (internal-NACC)",          na.omit(aggregate_all_NACC_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "aggregated-cox",
  "AIC feature selection (internal-NACC)",  na.omit(aggregate_fts_NACC_results1_MCI$AUC_list))

# time-cox(FADRS-MCI)
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "time-cox(FADRS-MCI)",
  "APOE4+age (internal-NACC)",             na.omit(time_varying_APOE4_NACC_full_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "time-cox(FADRS-MCI)",
  "full features (internal-NACC)",          na.omit(time_varying_all_NACC_results1_MCI$AUC_list))
boxplot_auc_index_df_MCI <- add_data(boxplot_auc_index_df_MCI, "time-cox(FADRS-MCI)",
  "AIC feature selection (internal-NACC)",  na.omit(time_varying_all_NACC_AIC_results1_MCI$AUC_list))

boxplot_auc_index_df_MCI <- boxplot_auc_index_df_MCI %>%
  mutate(
    model = factor(model, levels = c(
      "baseline-cox[init]", "baseline-cox[last]",
      "aggregated-cox", "time-cox(FADRS-MCI)"
    )),
    feature_set = factor(feature_set, levels = c(
      "APOE4+age (internal-NACC)",
      "full features (internal-NACC)",
      "AIC feature selection (internal-NACC)"
    ))
  ) %>%
  filter(is.finite(value))


panel_A_df_MCI <- bind_rows(
  # internal — APOE4+age
  boxplot_auc_index_df_MCI %>%
    filter(feature_set == "APOE4+age (internal-NACC)") %>%
    mutate(feature_set = "APOE4+age"),

  # internal — full features
  boxplot_auc_index_df_MCI %>%
    filter(feature_set == "full features (internal-NACC)") %>%
    mutate(feature_set = "full features"),

  # internal — AIC selected
  boxplot_auc_index_df_MCI %>%
    filter(feature_set == "AIC feature selection (internal-NACC)") %>%
    mutate(feature_set = "AIC selected"),

  # external — AIBL (all 4 models)
  data.frame(
    model       = "baseline-cox[init]",
    feature_set = "external (AIBL)",
    value       = na.omit(baseline_AIBL_results1_MCI$AUC_list)
  ),
  data.frame(
    model       = "baseline-cox[last]",
    feature_set = "external (AIBL)",
    value       = na.omit(baseline_AIBL_results2_MCI$AUC_list)
  ),
  data.frame(
    model       = "aggregated-cox",
    feature_set = "external (AIBL)",
    value       = na.omit(aggregate_fts_AIBL_results1_MCI$AUC_list)
  ),
  data.frame(
    model       = "time-cox(FADRS-MCI)",
    feature_set = "external (AIBL)",
    value       = na.omit(time_varying_external_results1_MCI$AUC_list)
  )
) %>%
  mutate(
    feature_set = factor(feature_set, levels = c(
      "APOE4+age", "full features", "AIC selected", "external (AIBL)"
    )),
    model = factor(model, levels = c(
      "baseline-cox[init]", "baseline-cox[last]",
      "aggregated-cox", "time-cox(FADRS-MCI)"
    ))
  ) %>%
  filter(is.finite(value))

p_A_MCI <- plot_boxplot_with_anno(
  df             = panel_A_df_MCI,
  y_label        = "AUC",
  save_path      = "../results/MCI_panel_A_auc_model_comparison.pdf",
  compare_group1 = "baseline-cox[init]",
  compare_group2 = "time-cox(FADRS-MCI)",
  fill_var       = "model"
)
print(p_A_MCI)
```

```{r}
# ============================================================
# PANEL B: AUC Boxplot — visit comparison, time-cox only
# x-axis: APOE4+age | full features | AIC selected | external (AIBL)
# fill: visit count
# ============================================================

panel_B_df_MCI <- bind_rows(
  # internal — APOE4+age, all visits
  data.frame(feature_set = "APOE4+age", visit = "full visits",
             value = na.omit(time_varying_APOE4_NACC_full_results1_MCI$AUC_list)),
  data.frame(feature_set = "APOE4+age", visit = "3-visits",
             value = na.omit(time_varying_APOE4_NACC_full_results2_MCI$AUC_list)),
  data.frame(feature_set = "APOE4+age", visit = "2-visits",
             value = na.omit(time_varying_APOE4_NACC_full_results3_MCI$AUC_list)),
  data.frame(feature_set = "APOE4+age", visit = "1-visit",
             value = na.omit(time_varying_APOE4_NACC_full_results4_MCI$AUC_list)),

  # internal — full features, all visits
  data.frame(feature_set = "full features", visit = "full visits",
             value = na.omit(time_varying_all_NACC_results1_MCI$AUC_list)),
  data.frame(feature_set = "full features", visit = "3-visits",
             value = na.omit(time_varying_all_NACC_results2_MCI$AUC_list)),
  data.frame(feature_set = "full features", visit = "2-visits",
             value = na.omit(time_varying_all_NACC_results3_MCI$AUC_list)),
  data.frame(feature_set = "full features", visit = "1-visit",
             value = na.omit(time_varying_all_NACC_results4_MCI$AUC_list)),

  # internal — AIC selected, all visits
  data.frame(feature_set = "AIC selected", visit = "full visits",
             value = na.omit(time_varying_all_NACC_AIC_results1_MCI$AUC_list)),
  data.frame(feature_set = "AIC selected", visit = "3-visits",
             value = na.omit(time_varying_all_NACC_AIC_results2_MCI$AUC_list)),
  data.frame(feature_set = "AIC selected", visit = "2-visits",
             value = na.omit(time_varying_all_NACC_AIC_results3_MCI$AUC_list)),
  data.frame(feature_set = "AIC selected", visit = "1-visit",
             value = na.omit(time_varying_all_NACC_AIC_results4_MCI$AUC_list)),

  # external — AIBL, time-cox only, all visits
  data.frame(feature_set = "external (AIBL)", visit = "full visits",
             value = na.omit(time_varying_external_results1_MCI$AUC_list)),
  data.frame(feature_set = "external (AIBL)", visit = "3-visits",
             value = na.omit(time_varying_external_results2_MCI$AUC_list)),
  data.frame(feature_set = "external (AIBL)", visit = "2-visits",
             value = na.omit(time_varying_external_results3_MCI$AUC_list)),
  data.frame(feature_set = "external (AIBL)", visit = "1-visit",
             value = na.omit(time_varying_external_results4_MCI$AUC_list))
) %>%
  mutate(
    feature_set = factor(feature_set, levels = c(
      "APOE4+age", "full features", "AIC selected", "external (AIBL)"
    )),
    visit = factor(visit, levels = c(
      "full visits", "3-visits", "2-visits", "1-visit"
    ))
  ) %>%
  filter(is.finite(value)) %>%
  rename(model = visit)    # rename so function works with fill_var

# rename back
panel_B_df_MCI <- panel_B_df_MCI %>% rename(visit = model)

p_B_MCI <- plot_boxplot_with_anno(
  df             = panel_B_df_MCI,
  y_label        = "AUC",
  save_path      = "../results/MCI_panel_B_auc_visit_comparison.pdf",
  compare_group1 = "full visits",
  compare_group2 = "1-visit",
  fill_var       = "visit"
)
print(p_B_MCI)
```

```{r}
print(final_selected_feature_MCI)
```


```{r}
# with feature renaming
create_forest_plot(
  summary_results = time_varying_external_results1_MCI$model_summary,
  title           = "Forest Plot — Time-Varying Cox (MCI)",
  file_path_name  = "../results/MCI_panel_D_forest_plot.pdf",
  feature_rename  = c(
    "CDR_SB"       = "CDR Sum of Boxes",
    "MMSE"         = "MMSE Score",
    "memory_recall2" = "LM-II",
    "animal"       = "Animal Fluency Test",
    # "GDS"          = "Geriatric Depression Scale",
    # "APOE4"        = "APOE ε4 Status",
    "age"          = "Age",
    "edu"          = "Education",
    "blood_pressure_diastolic" = "BP (diastolic)"
  )
)
```




####### clustering

```{r}
risk_group_train_MCI <- time_varying_external_results1_MCI$risk_train %>%
  mutate(risk_group = case_when(
    last_risk_score < quantile(last_risk_score, 0.5) ~ "Low Risk",
    # last_risk_score < quantile(last_risk_score, 0.66) ~ "Medium Risk",
    TRUE ~ "High Risk"
  ))
analyze_risk_groups(NACC_full_df_MCI, risk_group_train_MCI)
```


```{r}
risk_group_means_train_MCI <- sapply(unique(risk_group_train_MCI$risk_group), function(group) {
  group_ids <- risk_group_train_MCI$ID[risk_group_train_MCI$risk_group == group]
  subset_results <- time_varying_external_results1_MCI$risk_train[time_varying_external_results1_MCI$risk_train$ID %in% group_ids, ]
  mean(subset_results$last_risk_score, na.rm = TRUE)
})
names(risk_group_means_train_MCI) <- unique(risk_group_train_MCI$risk_group)
risk_group_means_train_MCI
```


```{r}
risk_group_test_MCI <- time_varying_external_results1_MCI$risk_test %>%
  mutate(
    risk_group = sapply(last_risk_score, function(score) {
      differences <- abs(risk_group_means_train_MCI - score)
      names(differences)[which.min(differences)]
    })
  )

analyze_risk_groups(AIBL_full_df_imputed_MCI, risk_group_test_MCI)
```

```{r}
# selected_features_trajectories_MCI <- c("CDR_SB", "memory_recall2","animal", "MMSE", "BMI")
# 
# centroid_trajectory_test_MCI = plot_centroid_trajectories(AIBL_full_df_imputed_MCI, risk_group_test_MCI, selected_features_trajectories_MCI, "../results/MCI feature_trajectories_AIBL.pdf")
```

```{r}
selected_features_trajectories_MCI = c("animal", "blood_pressure_diastolic","GDS", "memory_recall2", "MMSE")
centroid_trajectory_test_MCI <- plot_centroid_trajectories(
  df                             = AIBL_full_df_imputed_MCI,
  risk_group                     = risk_group_test_MCI,
  selected_features_trajectories = selected_features_trajectories_MCI,
  pdf_path                       = "../results/MCI_panel_E_feature_trajectories_AIBL.pdf",
  feature_rename = c(
    "memory_recall2" = "LM-II",
    "animal"         = "Animal Fluency Test",
    "MMSE"           = "MMSE Score",
    "blood_pressure_diastolic" = "BP (diastolic)"
    # "GDS"            = "Geriatric Depression Scale"
  ),
  palette = c("#9bbf8a", "#fa8878")   # High Risk, Low Risk
)
```


```{r}
risk_group_means_test_MCI <- sapply(unique(risk_group_test_MCI$risk_group), function(group) {
  group_ids <- risk_group_test_MCI$ID[risk_group_test_MCI$risk_group == group]
  subset_results <- time_varying_external_results1_MCI$risk_test[time_varying_external_results1_MCI$risk_test$ID %in% group_ids, ]
  mean(subset_results$last_risk_score, na.rm = TRUE)
})
names(risk_group_means_test_MCI) <- unique(risk_group_test_MCI$risk_group)
risk_group_means_test_MCI
```

```{r}
risk_group_test_MCI

```
